---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# zzmesoimpute

*`r format(Sys.time(), '%Y-%m-%d %H:%M %Z')`*

`zzmesoimpute` provides methods for imputing left-censored
(below-limit-of-detection) concentrations from multiplexed immunoassays
such as Meso Scale Discovery (MSD). Readings below an assay's detection
limit are not missing but left-censored: the true concentration is known
only to lie in the interval from zero to a per-assay detection limit. The
package implements a common interface across five imputation strategies so
that the incumbent laboratory convention can be benchmarked against
established censored-data methods.

The methods are evaluated in a companion manuscript compendium,
`zzmesoimputepaper`, which calibrates a simulation study to the censoring
structure of an MSD panel.

## Installation

```{r, eval = FALSE}
# install.packages("remotes")
remotes::install_github("rgt47/zzmesoimpute")
```

## Imputation strategies

All imputation functions share the interface `impute_*(x, lod, censored,
...)`, where `x` is the concentration vector, `lod` is a single detection
limit or one per observation, and `censored` is a logical vector marking
below-detection readings (defaulting to `is.na(x)`).

| Function | Strategy | Reference |
|----------|----------|-----------|
| `impute_substitution()` | Fixed fraction of the LOD (`fraction = 0.5` gives LOD/2; `1/sqrt(2)` gives LOD/sqrt(2)) | Hornung & Reed (1990) |
| `impute_uniform()` | Uniform(0, LOD) draw, the incumbent MSD lab method | -- |
| `impute_ros()` | Regression on order statistics | Helsel (2012) |
| `impute_mle()` | Censored-lognormal MLE conditional-mean imputation | Cohen (1959) |
| `impute_mi_lognormal()` | Proper multiple imputation under a censored lognormal | Rubin (1987) |

The single-imputation methods return a numeric vector the same length as
`x`. `impute_mi_lognormal()` returns a list of `m` completed vectors, to be
analysed separately and combined by Rubin's rules.

## Example

```{r example}
library(zzmesoimpute)

# Six readings; two fall below a detection limit of 1.5.
x        <- c(5.2, NA, 3.1, NA, 8.4, 2.7)
censored <- is.na(x)

# Single-imputation strategies return a numeric vector.
impute_substitution(x, lod = 1.5, censored = censored, fraction = 0.5)
impute_uniform(x, lod = 1.5, censored = censored, seed = 1)
impute_mle(x, lod = 1.5, censored = censored)

# Multiple imputation returns m completed datasets.
completions <- impute_mi_lognormal(x, lod = 1.5, censored = censored,
                                   m = 5, seed = 1)
length(completions)
completions[[1]]
```

## Reading MSD exports

`read_meso()` ingests a raw MSD plate-reader Excel export, locating the
header block and mapping the vendor columns (detection range, calculated
concentration, detection limit, exclusion flag) to tidy names suitable for
the imputation functions. See `meso_default_map()` for the column mapping
and `?read_meso` for details.

## License

GPL-3. See the companion manuscript for methodological background and a
simulation-based comparison of the strategies.
